The Astropy Project

### 2017-12-08 main.tex
intro: remove two non-sequitur sentences
The point of this paragraph is to briefly state the relevant history, and to describe the current state of astropy.

A general statement about aims would belong further up (I'm not sure where, so I did not put it anywhere). The object of "these tools" in original wording was not clear.
Edits for Section 3.6

\python\footnote{\url{https://www.python.org/}} is an increasingly popular, general-purpose
programming language that is available under a permissive open source software license free of
charge for all major operating systems. The programming language has become especially popular
in the quantitative sciences, where researchers must simultaneously produce research, perform
data analysis, and develop software. A large part of this success owes itself to the vibrant
community of developers and a continuously-growing ecosystem of tools, web services, and stable
well-developed packages that enable easier collaboration on software development, easier
writing and sharing of software documentation, and continuous testing and validation of
software. While dedicated libraries provide support for array representation and
arithmetic \citep[\package{numpy};][]{numpy}, a wide variety of functions for scientific
computing \citep[\package{scipy};][]{scipy}, and publication-quality plotting
\citep[\package{matplotlib};][]{matplotlib}, tens of thousands of other high-quality and easy-to-use
packages are available, which can help with tasks that are not astronomy specific but
might be performed in the course of astronomical research; e.g., interfacing with
databases or statistical inferences. More recently, the development of package managers such as
Anaconda\footnote{\url{https://anaconda.org/}} have streamlined the installation process for
most packages, significantly reducing the entry barrier for using many such libraries.

@@

The \package{astropy.io} subpackage provides support for reading and writing
data to a variety of ASCII and binary file formats, such as a wide range of
ASCII data table formats, FITS, and VOTable.
It also provides a unified interface for reading and writing data with these
different formats using the \package{astropy.table} subpackage.
For many common cases this simplifies the process of file input and output and
reduces the need to master the separate details of all the I/O packages within
\astropypkg.

@@

The \package{astropy.io.ascii} subpackage now includes a significantly faster
Cython/C engine for reading and writing ASCII files. This is available for the
most common formats.  On average the new engine is about 4 to 5 times faster
than the corresponding pure-\python implementation, and is often comparable to
the speed of the \package{Pandas} \citep{pandas} ASCII file interface.  The
fast reader has parallel processing option that allows harnessing multiple
cores for input parsing to achieve even greater speed gains.  By default,
\texttt{read()} and \texttt{write()} will attempt to use the fast C engine
when dealing with compatible formats. Certain features of the full read
/ write interface are not available in the fast version, in which case the
pure-\python version will automatically be used.

@@

The \package{astropy.io.ascii} subpackage now provides the capability
to read a table within an HTML file or web URL into an astropy
\texttt{Table} object. Conversely a \texttt{Table} object can now
be written out as an HTML table.

@@

The \package{astropy.io.fits} subpackage started as a direct port of the
PyFITS project \citep{PyFITS}. Therefore it is pretty stable, with mostly bug
fixes but also a few new features and performance improvements.  The API
remains compatible with PyFITS, which is now deprecated in favor of
\astropypkg.

Command-line scripts are now available for printing a summary of the HDUs in
a FITS file(s) (\texttt{fitsinfo}) and for printing the header information to
the screen in a human-readable format (\texttt{fitsheader}).

FITS files are now loaded \emph{lazily} (the default behavior, which can be
deactivated if needed), where all HDUs are not loaded until they are
requested. This should provide substantial speedups for situations using the
convenience functions (e.g., \texttt{getheader()} or \texttt{getdata()}) to
get HDU’s that are near the front of a file with many HDUs.

>>>>

\python\footnote{\url{https://www.python.org/}} is an increasingly popular, general-purpose
programming language that is available under a permissive open source software license free of
charge for all major operating systems. The programming language has become especially popular
in the quantitative sciences, where researchers must simultaneously produce research, perform
data analysis, and develop software. A large part of this success owes itself to the vibrant
community of developers and a continuously-growing ecosystem of tools, web services, and stable
well-developed packages that enable easier collaboration on software development, easier
writing and sharing of software documentation, and continuous testing and validation of
software. While dedicated libraries provide support for array representation and
arithmetic \citep[\package{numpy};][]{numpy}, a wide variety of functions for scientific
computing \citep[\package{scipy};][]{scipy}, and publication-quality plotting
\citep[\package{matplotlib};][]{matplotlib}, tens of thousands of other high-quality and easy-to-use
packages are available, which can help with tasks that are not astronomy specific but
might be performed in the course of astronomical research; e.g., interfacing with
databases or statistical inferences. More recently, the development of package managers such as
Anaconda\footnote{\url{https://anaconda.org/}} has significantly streamlined the installation
process for many libraries, lowering many barriers to entry.

@@

The \package{astropy.io} subpackage provides support for reading and writing
data to a variety of ASCII and binary file formats, such as a wide range of
ASCII data table formats, FITS, and VOTable.
It also provides a unified interface for reading and writing data with these
different formats using the \package{astropy.table} subpackage.
For many common cases, this simplifies the process of file input and output (I/O) and
reduces the need to master the separate details of all the I/O packages within
\astropypkg.

@@

The \package{astropy.io.ascii} subpackage now includes a significantly faster
Cython/C engine for reading and writing ASCII files. This is available for most
of the common formats.  On average, the new engine is about 4 to 5 times faster
than the corresponding pure-\python implementation and is often comparable to
the speed of the \package{pandas} \citep{pandas} ASCII file interface.  The
fast reader has parallel processing option that allows harnessing multiple
cores for input parsing to achieve even greater speed gains.  By default,
\texttt{read()} and \texttt{write()} will attempt to use the fast Cython/C engine
when dealing with compatible formats. Certain features of the full read
/ write interface are unavailable in the fast version, in which case the
pure-\python version will automatically be used.

@@

The \package{astropy.io.ascii} subpackage now provides the capability
to read a table within an HTML file or web URL into an \astropypkg
\texttt{Table} object. Conversely, a \texttt{Table} object can now
be written out as an HTML table.

@@

The \package{astropy.io.fits} subpackage started as a direct port of the
PyFITS project \citep{PyFITS}. Therefore, it is pretty stable, with mostly bug
fixes but also a few new features and performance improvements.  The API
remains mostly compatible with PyFITS, which is now deprecated in favor of
\astropypkg.

Command-line scripts are now available for printing a summary of the HDUs in
FITS file(s) (\texttt{fitsinfo}) and for printing the header information to
the screen in a human-readable format (\texttt{fitsheader}).

FITS files are now loaded \emph{lazily} (this default behavior can be
deactivated, if needed), where all HDUs are not loaded until
requested. This should provide substantial speed-ups when using the
convenience functions (e.g., \texttt{getheader()} or \texttt{getdata()}) to
get an HDU that is near the front of many in a file.
